---
title: "CrimeMD"
output: html_document
---

## Reading in the Data
We load up all the libaries we plan on using for this dataset and load in the crimemd csv from my files. 

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(DT)
library(ggrepel)
library(leaflet)
library(janitor)
library(MASS)
library(scales)
library(lubridate)
setwd("C:/Users/Jeffo/Downloads")
crimemd<- read.csv('crimemd.csv')
crimemd <- crimemd%>%
  clean_names()
  
```

## Cleaning the Data
Before we start to anaylze the data, we need to clean the data first in order have to data to be more user friendly. 

### Renaming some variables for easier reading

```{r}
colnames(crimemd)[36] <- "b_e_percent_change"
colnames(crimemd)[20] <- "overall_percent_change"
colnames(crimemd)[22] <- "violent_crime_rate_change"
```
### using the lubridate package, get date and year
```{r}
date<-as.Date(crimemd$year,'%m/%d/%Y')
year<-as.numeric(format(date,'%Y'))
crimemd$date<- date
crimemd$year<- year
```
## Exploratory Anaylsis
```{r}
sort(crimemd$population)
```

### 5 Number summary for population
Using the 5 number summary, we are able to see that the Minimum population is 4098000. The Maximum population is 6016447
```{r}
summary(crimemd$population)
```

##  Boxplot for Population
I made a boxplot for population in order to see the population statstics a bit easier. 
```{r}
par(mfrow = c(1, 2))
boxplot(crimemd$population)
boxplot(crimemd$population)


abline(h = min(crimemd$population), col = "Blue")

abline(h = max(crimemd$population), col = "Yellow")
       
abline(h = median(crimemd$population), col = "Green")

abline(h = quantile(crimemd$population, c(0.25, 0.75)), col = "Red")
```

## Focusing on Breaking and Entering
Since our project is mainly focusing on burglary, breaking and entering is the closest there is to that. 
```{r}
sort(crimemd$b_e)
```

## 5 number summary for Breaking and Entering
We can see that the minumum for breaking and entering occurences is 24822. The max is 71130. 
```{r}
summary(crimemd$b_e)
```

## Boxplot for Breaking and Entering
```{r}
par(mfrow = c(1, 2))
boxplot(crimemd$b_e)
boxplot(crimemd$b_e)


abline(h = min(crimemd$b_e), col = "Blue")

abline(h = max(crimemd$b_e), col = "Yellow")
       
abline(h = median(crimemd$b_e), col = "Green")

abline(h = quantile(crimemd$b_e, c(0.25, 0.75)), col = "Red")
```

## Line Graph for MD crimes
We can see that overall crimes in MD have gone steadily down since 1995, and are currently at its lowest in 40 years. 
```{r}
plot<-ggplot(data=crimemd, aes(x=year, y=grand_total))+geom_line(color = 'orange')+labs(y="Crimminal Occurences", title=" Crimes in MD from 1970-2019")+theme(plot.title = element_text(hjust =0.5))
plot+theme(legend.position = "none")
```

## Line graph for overall % change in crimes
The sharp decrease in crimes in MD from 1980-1983 is seen through this graph. Almost a 20 percent difference in 5 years. 
```{r}
plot<-ggplot(data=crimemd, aes(x=year, y=overall_percent_change))+geom_line(color = 'red')+labs(y="% change ", title="Percent Change in overall crimes in MD from 1970-2019")+theme(plot.title = element_text(hjust =0.5))
plot+theme(legend.position = "none")
```

## Line Graph for breaking and Entering
We can see that there was a sharp decrease in breaking and entering crimes in MD from 1980-1983. 
```{r}
plot<-ggplot(data=crimemd, aes(x=year, y=b_e))+geom_line(color = 'red')+labs(y="Breaking and Entering Occurences", title="Breaking and Entering Crimes in MD from 1970-2019")+theme(plot.title = element_text(hjust =0.5))
plot+theme(legend.position = "none")
```

## Line graph for percent change for breaking and entering
We can see that huge percentage change from 1980-1983 which leads me to believe a law or jurisdiction must have affected this. The same as well with 2000-2002 when there was that big percent increase. 
```{r}
plot<-ggplot(data=crimemd, aes(x=year, y=b_e_percent_change))+geom_line(color = 'red')+labs(y="Breaking and Entering % change ", title="Breaking and Entering Percent Change in MD from 1970-2019")+theme(plot.title = element_text(hjust =0.5))
plot+theme(legend.position = "none")
```

